---
title: "Blog 5: Personality and Worldview"
author: "April Zheng"
date: "October 6, 2025"
output:
  pdf_document: default
  word_document: default
urlcolor: blue
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(readr) # read_csv
library(ggplot2) # ggplot
library(tidyr)
library(gridExtra) # grid.arrange
library(psych) #fa.parallel
```

# Measuring Personality

Carney et al. (2009) use several strategies to assess personality in their investigation of the link between personality and political ideology. This week's assignment will help you understand aspects of these methodologies and associated challenges.

The survey you took included two different short versions of the Big Five Inventory. You will have the chance to compare these two inventories and to get some experience with the Personal Living Space Cue Inventory (PLSCI) ([Gosling et al. 2005](http://gosling.psy.utexas.edu/wp-content/uploads/2014/09/EnvandBehav05GoslingetalPLSCI.pdf)), which was also used in Carney et al. (2009).  Some of your answers will be recorded in a new survey that you will receive in your email.

*To receive credit for this blog post, you must complete questions 1-4, and at least one additional question.*

**Data Details:**

* File Name 1: \texttt{photo\_assignment.csv} These are the photo evaluation assignments for questions 1-4.

* File Name 2: \texttt{Oct6ClassData.csv}

* Source: These data are from the survey you completed at home. You took the BFI-10, a short version of the Big Five Inventory ([Rammstedt and John 2007](https://psycnet.apa.org/record/2007-02371-015)) and the Ten-Item Personality Inventory (TIPI), which is a different instrument designed to quickly meausre the Big Five ([Gosling et al. 2003](http://gosling.psy.utexas.edu/wp-content/uploads/2014/09/JRP-03-tipi.pdf)). You then answered the same questions used by Carney et al. (2009) to assess political attitudes in some of their studies.

Variable Name         | Variable Description
--------------------- | --------------------------------------
\texttt{Overall}         | Self-reported overall ideology on a 1-5 scale with 1 being extremely liberal and 5 being extremely conservative
\texttt{Social}       | Self-reported social ideology on a 1-5 scale with 1 being extremely liberal and 5 being extremely conservative
\texttt{Economic}    | Self-reported economic ideology on a 1-5 scale with 1 being extremely liberal and 5 being extremely conservative
\texttt{Random ID}| A randomly generated respondent identifier
\texttt{BFI\_extraversion}        | The average of the two items on the BFI-10 associated with extraversion
\texttt{BFI\_agreeableness}    | The average of the two items on the BFI-10 associated with agreeableness
\texttt{BFI\_conscientiousness}| The average of the two items on the BFI-10 associated with conscientiousness
\texttt{BFI\_emot\_stability}| The average of the two items on the BFI-10 associated with emotional stability (also known as neuroticism, although low emotional stability is the same as high neuroticism)
\texttt{BFI\_openness} | The average of the two items on the BFI-10 associated with openness
\texttt{TIPI\_extraversion}| The average of the two items on the TIPI-10 associated with extraversion
\texttt{TIPI\_agreeableness}    | The average of the two items on the TIPI-10 associated with agreeableness
\texttt{TIPI\_conscientiousness}| The average of the two items on the TIPI-10 associated with conscientiousness
\texttt{TIPI\_emot\_stability}| The average of the two items on the TIPI-10 associated with emotional stability(also known as neuroticism)
\texttt{TIPI\_openness} | The average of the two items on the TIPI-10 associated with openness

The last 20 variables in the data are the actual items from the BFI-10 and the TIPI. If you want to refresh your memory about the questions and learn which questions were meant to go together, you can find more about the BFI-10 [here](https://www.ocf.berkeley.edu/~johnlab/bfi.htm) (scroll down to the "Is there a shorter version of the BFI available?" question) and the TIPI [here](http://gosling.psy.utexas.edu/scales-weve-developed/ten-item-personality-measure-tipi/ten-item-personality-inventory-tipi/).

Don't forget to load the data.

```{r faces, include=FALSE}
personality_data <- read_csv('Oct6ClassData.csv')
```

## Question 1
Check the Google Drive for your photo assignments and download the assigned photos of living space (you should be assigned to three photos). **Use the worksheet on the Google Drive to do an assessment based on the PLSCI. Based on the PLSCI and your own general impressions, fill out the new class survey (in your email) with your guess about each person's openness, conscientiousness, and overall political ideology without looking at the actual data. Make sure to write down your responses to the survey since you won't have access to them after you submit. (Page 391 of [Gosling et al. 2002](http://gosling.psy.utexas.edu/wp-content/uploads/2014/09/JPSP02-Roomwithacue.pdf) might help you figure out which items on the inventory will be most helpful here.) Use the random ID to check your guesses. How did you do?** (If you are finding the PLSCI difficult and frustrating, peek ahead to the next few questions.)

See Google Doc

## Question 2
Each photo should have at least three people assigned to code it. Confer with each of the people who were also assigned to your photo and see what their ratings were on the PLSCI, as well as their guesses about openness, conscientiousness, and ideology. **What does this tell you about the PLSCI? Read a bit about intercoder (aka inter-rater) reliability and reflect on its importance in research like this.**

See Google Doc

## Question 3
**What challenges did you encounter administering the PLSCI? What do you notice that might complicate applying this inventory today?**

See Google Doc

## Question 4
**How would you redesign the PLSCI to make it more useful/current?**

See Google Doc

## Question 5
Let's see how the class compares to a large dataset of people who have taken one of these personality inventories. [Gosling et al. (2003)](https://citeseerx.ist.psu.edu/viewdoc/download?doi=10.1.1.113.6704&rep=rep1&type=pdf) report findings from administering the TIPI to about 1800 undergraduates. **Compare the class results to the published norms (look at the table on page 526, all ethnicities, Whole sample; i.e. the upper left corner of the table). Make a table or a plot of the class results and then write your thoughts about how these are similar or different to international norms and why that might be the case.**
```{r}
personality_data |> select(10:14) |> sapply(mean, na.rm=TRUE)
#sapply applies a function across a vector (vector = column) 
#Here we are finding the mean for all the columns of BFI and TIPI. 

personality_data |> select(10:14) |> sapply(sd, na.rm=TRUE)

```

